Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-25 paper/paper.tex
Fix typo and add comment.

Using this model, we may view the problem of estimating the similarity $Y$ of
items $A$ and $B$ that most likely produced ratings $R_A$ and $_B$ as
that of finding the Maximum Likelihood Estimate of $Y | \vec{R}$:
\begin{align}
P(Y | \vec{R}) = \frac{P(\vec{R} | Y)P(Y)}{P(\vec{R})}.
\end{align}

\begin{align}
\hat{y} &= \argmax_yP(Y=y|\vec{R}=\vec{r}) 
\\&= \argmax_y\left[\frac{P(\vec{R}=\vec{r} |
Y=y)P(Y=y)}{P(\vec{R}=\vec{r})}\right]
\\&= \argmax_y\left[P(\vec{R}=\vec{r}|Y=y)P(Y=y)\right]
\\&=
\argmax_y\left[\frac{1}{2\pi\sqrt{\left|\mathbf{\Sigma_y}\right|}}\exp{\left(-\frac{1}{2}
(\vec{r} - \vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r})\right)}
\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
{2\sigma_{s}^2}\right)}\right]
\\&= \argmin_y\left[(\vec{r} -
\vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r}) + \frac{(y -
\mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&=
\argmin_y\left[\vec{r^\prime}^\mathrm{T}\mathbf{\Sigma_y}^{-1}\vec{r^\prime}
 + \frac{(y - \mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&= \argmin_y\left[\frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right],
\end{align}
where we used the fact that
\begin{align}
\mathbf{\Sigma_y}^{-1} = \frac{1}{\sigma_r^2}
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]^{-1} =
\frac{1}{\sigma_r^2}\frac{1}{1 - y^2}\left[ \begin{array}{cc}  1 & -y \\ -y & 1
\end{array} \right].
\end{align}
and that
\begin{align}
\left|\mathbf{\Sigma}\right| = \sigma_r^2(1 - y^2)
\end{align}
along with the substitution $\vec{r^\prime} = (r_A^\prime~~r_B^\prime) = \vec{r}
- \vec{\mu_r}$. Next, we find the minimum by taking the derivative with respect
to $y$ and setting it to zero:
\begin{align}
\frac{d}{dy}\left[ \frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\
-r_A^\prime r_B^\prime y^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)y - r_A^\prime
r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - y^2)^2(y - \mu_s) - \sigma_r^2(1
- y^2)y &= 0.
\end{align}
This final expression implicitly yields the MLE of $y$ as a function of the
ratings $r_A$ and $r_B$. That is, constraining the parameters $\mu_r$,
$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values, this expression defines
a surface that can be to efficiently estimate $y$ given a pair of ratings.

>>>>

Using this model, we may view the problem of estimating the similarity $Y$ of
items $A$ and $B$ that most likely produced ratings $R_A$ and $R_B$ as
that of finding the Maximum Likelihood Estimate of $Y | \vec{R}$:
\begin{align}
P(Y | \vec{R}) = \frac{P(\vec{R} | Y)P(Y)}{P(\vec{R})}.
\end{align}

\begin{align}
\hat{y} &= \argmax_yP(Y=y|\vec{R}=\vec{r}) 
\\&= \argmax_y\left[\frac{P(\vec{R}=\vec{r} |
Y=y)P(Y=y)}{P(\vec{R}=\vec{r})}\right]
\\&= \argmax_y\left[P(\vec{R}=\vec{r}|Y=y)P(Y=y)\right]
\\&=
\argmax_y\left[\frac{1}{2\pi\sqrt{\left|\mathbf{\Sigma_y}\right|}}\exp{\left(-\frac{1}{2}
(\vec{r} - \vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r})\right)}
\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
{2\sigma_{s}^2}\right)}\right]
\\&= \argmin_y\left[(\vec{r} -
\vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r}) + \frac{(y -
\mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&=
\argmin_y\left[\vec{r^\prime}^\mathrm{T}\mathbf{\Sigma_y}^{-1}\vec{r^\prime}
 + \frac{(y - \mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&= \argmin_y\left[\frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right],
\end{align}
where we used the fact that
\begin{align}
\mathbf{\Sigma_y}^{-1} = \frac{1}{\sigma_r^2}
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]^{-1} =
\frac{1}{\sigma_r^2}\frac{1}{1 - y^2}\left[ \begin{array}{cc}  1 & -y \\ -y & 1
\end{array} \right].
\end{align}
and that
\begin{align}
\left|\mathbf{\Sigma}\right| = \sigma_r^2(1 - y^2)
\end{align}
along with the substitution $\vec{r^\prime} = (r_A^\prime~~r_B^\prime) = \vec{r}
- \vec{\mu_r}$. Next, we find the minimum by taking the derivative with respect
to $y$ and setting it to zero:
\begin{align}
\frac{d}{dy}\left[ \frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\
-r_A^\prime r_B^\prime y^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)y - r_A^\prime
r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - y^2)^2(y - \mu_s) - \sigma_r^2(1
- y^2)y &= 0.
\end{align}
This final expression implicitly yields the MLE of $y$ as a function of the
ratings $r_A$ and $r_B$. That is, constraining the parameters $\mu_r$,
$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values (e.g. $\sigma_r,
\sigma_s > 0$), this expression defines a surface that can be to efficiently
estimate $y$ given a pair of ratings.
